Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-04-18 priority-queues.tex
sequential executions

An \emph{execution} $e$ is a sequence of call and return actions which satisfy the following well-formedness properties: each return is preceded by a matching call (i.e., having the same operation identifier), and each operation identifier is used in at most one call/return. We assume every set of executions is closed under isomorphic renaming of operation identifiers. An execution is called \emph{sequential} when no two operations overlap, i.e., each call action is immediately followed by its matching return action, and \emph{concurrent} otherwise. To ease the reading, we write a sequential execution as a sequence of $\textit{put}(a,p)$ and $\textit{rm}(a)$ symbols representing a pair of actions $\textit{call}_o(\textit{put},a,p)\cdot \textit{ret}_o(\textit{put},a,p)$ and $\textit{call}_o(\textit{rm},a)\cdot \textit{ret}_o(\textit{rm},a)$, respectively (where $o\in\mathbb{O}$). For example, given two priorities $p_1 \prec p_2$, $\textit{put}(a,p_2) \cdot \textit{put}(b,p_1) \cdot \textit{rm}(b)$ is a sequential execution of the priority queue ($\textit{rm}$ returns $b$ because it has smaller priority). 
The set of all the possible sequential executions (obtained by adding and removing arbitrarily values with arbitrary priorities) is denoted by $\seqPQ$. TODO ADD THE REFERENCE TO THE LTS IN THE APPENDIX

@@

Linearizability \cite{journals/toplas/HerlihyW90} is the \emph{de facto} standard correctness condition for concurrent data structures. Given concurrent execution $e$ and sequential execution $s$ of length $n$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if they have same set of operations, and there is a bijection $f$ between operations s.t.

>>>>

An \emph{execution} $e$ is a sequence of call and return actions which satisfy the following well-formedness properties: each return is preceded by a matching call (i.e., having the same operation identifier), and each operation identifier is used in at most one call/return. We assume every set of executions is closed under isomorphic renaming of operation identifiers. An execution is called \emph{sequential} when no two operations overlap, i.e., each call action is immediately followed by its matching return action, and \emph{concurrent} otherwise. To ease the reading, we write a sequential execution as a sequence of $\textit{put}(a,p)$ and $\textit{rm}(a)$ symbols representing a pair of actions $\textit{call}_o(\textit{put},a,p)\cdot \textit{ret}_o(\textit{put},a,p)$ and $\textit{call}_o(\textit{rm},a)\cdot \textit{ret}_o(\textit{rm},a)$, respectively (where $o\in\mathbb{O}$). For example, given two priorities $p_1 \prec p_2$, $\textit{put}(a,p_2) \cdot \textit{put}(b,p_1) \cdot \textit{rm}(b)$ is a sequential execution of the priority queue ($\textit{rm}$ returns $b$ because it has smaller priority).

We define $\seqPQ$, the set of sequential priority queue executions semantically via labelled transition system ($LTS$). An LTS is a tuple $(Q,\Sigma,\rightarrow,q_0)$, where $Q$ is a set of states, $\Sigma$ is an alphabet of transition labels, $\rightarrow\subseteq Q\times\Sigma\times Q$ is a transition relation and $q_0$ is the initial state. We modelled priority queue as an LTS $\textit{LTS}_e$ as follows: Each state of $\textit{LTS}_e$ is a function from $\mathbb{P}$ into a finite sequence over $\mathbb{D}$ and represents a snapshot of priority queue content; The transition label are $\textit{put}$ and $\textit{rm}$ operations; Each transition modified state according to operations. For example, $q_1 \xrightarrow{\textit{rm}(\textit{empty})} q_2$, if $q_1 = q_2$, and they maps each element in $\mathbb{P}$ into $\epsilon$. Let $\seqPQ$ be the set of traces of $\textit{LTS}_e$.

@@

Linearizability \cite{journals/toplas/HerlihyW90} is the \emph{de facto} standard correctness condition for concurrent data structures. Given concurrent execution $e$ and sequential execution $s$ of length $n$, we say that $e$ is linearizable w.r.t $s$, denoted $e \sqsubseteq s$, if there is a bijection $f: O_1 \rightarrow O_2$, where $O_1$ and $O_2$ are set of operations of $e$ and $s$, respectively, and $f$ s.t.
